---
title: "Importance of Child's College Education"
author: "Joanna Choi"
description: "Differing views on the importance of child's college education based on parent's religion"
output:
  distill::distill_article:
    self_contained: false
---

  
## Introduction
  I am exploring: Does parents' religious beliefs in Christianity affect how important they view their child receiving a college education? My hypothesis is that Christian parents will want their children to attend college and receive a college education more than non-religious parents because the bible stresses the pursuit of knowledge and working hard, which relates to the traditional belief that children should pursue higher education.

  Here is the source I will use for this exploration:
[Pew Research Center - 2019 Survey of Religion Among U.S. Teens and Their Parents](https://www.pewresearch.org/religion/dataset/2019-survey-of-religion-among-u-s-teens-and-their-parents/). This data contains the information and variables I will need for this research including: the religious status (Christian or Non-Christian among the parents) which is the explanatory variable of interest (PBORN) and it is measured by whether or not the parent is a Christian (PBORN = 1) or not a Christian (PBORN = 2). Using the information of whether the parents are Christian or not, I will determine the importance they place on their child receiving a college education, which would be the outcome variable. The outcome variable (KIDTRAIT 1a) is measured on a scale where lower values mean that the parent thinks it is more important for their child to go to college whereas higher values mean that the parent thinks that it is not as important for their child to attend college. If I observe the pattern that the parents that are Christian are more likely to think that it is more important for their child to pursue a college education, this would support my hypothesis that the Christian beliefs are a factor that plays into these observed results, relating to a positive coefficient or correlation. However, if I observe that Christian parents view the importance of their child's college education equally or less seen than non-Christian parents, this would disprove my hypothesis, relating to a neutral or negative coefficient/relationship. Additionally, I will look at another variable, of how important parents view religion in their lives (PRELIMP) that is measured on a scale of 1 being very important to 4 being not at all important, as another factor to how important parents view their child going to college. I will look at this variable in conjunction with the other explanatory variable (PBORN) to conduct a multiple regression.

## Data/Design

```{r setup, include=FALSE}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
install.packages("haven", repos = "http://cran.us.r-project.org")
library("haven")
install.packages("modelsummary", repos = "http://cran.us.r-project.org")
```

```{r}
religion <- read_sav("data/religion/religionsurvey.sav")

religion

religion <- religion |>
  mutate(religious = case_when(PBORN == 1 ~ "Christian",
                               PBORN == 2 ~ "Not Christian")) |>
  drop_na(PBORN) 

educationplot <- religion |>
  drop_na(PBORN) |>
  ggplot(mapping = aes(x = religious, y = KIDTRAIT1b, na.rm = TRUE)) +
  geom_col() +
  labs(x = "Religious Belief", y = "Importance of Child's College Education")

educationplot
```
 This research is an observational study because this information/data is a result of a survey in which I was not able to actually conduct an experiment by assigning individuals to certain groups. Explanatory variables such as religion is not something I can assign someone or a group to and therefore, I am conducting research deriving from a "natural" variable coming from the individuals themselves rather than something I manipulated or treated. However, due to this factor, there is less internal validity because I was not able to control or account for the circumstances, conditions, or possible confounding variables between the Christian parents and non-Christian parents and therefore difficult to determine causal effect. However, on the other hand, this design allows me to have greater external validity in which this survey the data comes from, covers a larger population. The specific research design in this case would be a cross-sectional design in which I am using data acquired through observing/surveying the US population in 2019 for my analysis.
 
  Looking at the graph above, I organized the summarized my dependent variable of how important the parent views their child's college education, in relation to just the explanatory variable of whether the parent is a Christian or not. This scale of "importance of child's college education" is smaller when it is very important and as you get higher on the scale, it is less important. Therefore, taking this into account, it is observed that there is more importance placed on their child's college education if the parent is Christian compared to parents who are not Christian. However, the difference observed in the graph is very small and therefore finding the coefficient/finding statistical significance is important in searching for the relationship and there may also be other confounding variables that need to be taken into account, such as how important the parent views religion.
  
  
## Results
```{r}
library(broom)

coefficients1 <- lm(KIDTRAIT1b ~ PBORN, data = religion)

coefficients1 |>
  broom::tidy() |>
  select(term, estimate) |>  
  knitr::kable(digits = 2)
```
Looking at the table, the PBORN variable coefficient is -0.01. The coefficient is a negative value/slope and I can determine that there is a negative correlation/relationship between the parent being Christian and how important their child's college education is where the importance of their child going to college increases when the parent is Christian. However, even then, the coefficient is very small.

```{r}
education_importance <- religion |>
  drop_na(PBORN) |>
  mutate(`Religious Belief` = case_when(PBORN == 1 ~ "Christian",
                                        PBORN == 2 ~ "Not Christian")) |>
  ggplot(mapping = aes(x = PRELIMP, y = KIDTRAIT1b, color = `Religious Belief`)) +
  geom_point() +
  labs(x = "Importance of Religion", y = "Importance of Child's College Education")

education_importance
```
```{r}
coefficients2 <- lm(KIDTRAIT1b ~ PBORN + PRELIMP, data = religion)

coefficients2 |>
  broom::tidy() |>
  select(term, estimate) |>  
  knitr::kable(digits = 2)
```
```{r}
coefficients2 <- lm(KIDTRAIT1b ~ PBORN + PRELIMP, data = religion)

modelsummary::modelsummary(coefficients2,
                           statistic = c("p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```
  The main coefficient of interest was the estimate for PBORN which is -0.01. Because the coefficient of PRELIMP was 0, it shows that there exists a neutral relationship/association between the importance of religion and how important a parent views their child's college education (as also observed in the scatter plot). This may be because if a parent is not Christian, they are also not likely to not view religion as important in one's life, however this also does not account for the other religions. Nonetheless, holding this variable of PRELIMP constant, because the neutral association has already been established, we observe no change in the PBORN coefficient as it was -0.01 in the other regression as well. Therefore, it implies the negative association between the parent being Christian and how important their child going to college is once again, where the importance of their child going to college increases when the parent is Christian.
  
  This estimated coefficient for PBORN is not statistically significant. I found the p-value of PBORN to be p = 0.667 which is greater than the significance level when alpha equals 0.05. Therefore, given all these factors, I would not determine that the coefficient signifies a causal effect. The difference observed between the importance of a child's college education is very small between parents who are Christian and not Christian and there is no statistical significance observed for the PBORN coefficient as seen above. There may be also other confounding variables that may be accounting for the small difference observed and therefore, there is not enough evidence to determine a causal effect for this coefficient of PBORN.


## Conclusion
  I have observed that Christian parents are slightly more likely to vew their child's college education as more important than non-Christian parents but that difference is very marginal. Furthermore, after running multiple regression tests, I have observed that there is no statistical significance to signify a strong relationship/correlation between whether or not a parent is Christian and how important they view their child going to college. Some limitations of the study is that this data is only from 2019 which limits the data I was able to work with. The results may have differed if I was able to work with a larger range of years but that is something I can only determine through a future study. Additionally, the PBORN and PRELIMP variables are very similar in nature because PBORN records whether a parent is Christian or not and PRELIMP records how important a parent views religion. Therefore, in the future, I would try to use a different variable that has less possible correlation/similarity to the PBORN variable such as whether or not the parent went to college. 